The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-11 rational.tex
some work on constructions

\noindent 
\textit{Example.} The following combinator is represented by the given
diagram:

{{c : b * (1+1) <-> b + b}}

@@

% \paragraph*{Time Traveling.}
%% <<<<<<< local
%% A \emph{time traveling} intuition is also applicable here.  In a
%% normal cicruit, as computational steps are taken values flow from left
%% to right, i.e. computational time progresses from left to right.  The
%% action of {{eps+}} causes values, aka information particles, to flow
%% from right to left i.e. backwards in computational time.  The
%% interesting thing about this interpretation is that particles that
%% travel backwards in time get to see and interact with the history of
%% particles that they coexist with. This gives us an intuitive
%% interpretation of the isomorphism {{(-b1)*b2<->-(b1*b2)}} (see
%% Sec. \ref{sec:specific-constructions}) where it would otherwise seem
%% that {{-b1}} and {{b2}} move in opposite directions. The backward flow
%% of the {{-b1}} allows it to `see the past' of {{b2}} and is thus
%% equivalent to both particles moving backward in time.
%% =======
% A time traveling intuition is also applicable here: if view the
% left-to-right direction of information flow as the canonical forward
% direction of information, then the action of {{eps+}} allows values,
% aka information particles, to flow backwards in time. Particles that
% flow backwards in time get to see and interact with the history of
% particles that they coexist with, i.e., values that they exist
% multiplicatively with. Hence the isomorphism {{(-b1)*b2<->-(b1*b2)}}
% (whose construction is shown in Section \ref{sec:neg-constructions}).
%% >>>>>>> other

@@

As with additive duals, we represent {{eta*}} and {{eps*}} using
U-shaped connectors. In later diagrams, the wires for {{1}} will be
dropped as they can freely introduced or eliminated using {{uniti}}
and {{unite}}.

@@

%% {{(-b1)*b2 + b1*b2 <-> 0}}

@@

\begin{center}
\scalebox{0.8}{
  \includegraphics{diagrams/eps_fst.pdf}
}
\end{center}

@@

@@

@@

\paragraph*{QC} 
Are constructions are very similar to those developed in the context
of QC. Is this how nature computes? Are the analogies to time
traveling particles and anti-particles more concrete in some way?

>>>>

\noindent 
\textit{Example.}  We use the type {{bool}} as a shorthand to denote
the type {{1+1}} and use {{left ()}} to be {{true}} and {{right ()}}
to be {{false}}. The following combinator is represented by the given
diagram:

{{c : b * bool <-> b + b}}

@@

% \paragraph*{Time Traveling.}
%% <<<<<<< local
%% =======
% A time traveling intuition is also applicable here: if view the
% left-to-right direction of information flow as the canonical forward
% direction of information, then the action of {{eps+}} allows values,
% aka information particles, to flow backwards in time. Particles that
% flow backwards in time get to see and interact with the history of
% particles that they coexist with, i.e., values that they exist
% multiplicatively with. Hence the isomorphism {{(-b1)*b2<->-(b1*b2)}}
% (whose construction is shown in Section \ref{sec:neg-constructions}).
%% >>>>>>> other

@@

For convenience we sometime use the notation {{b1/b2}} to indicate the
pair {{b1 * (1/b2)}}.  As with additive duals, we represent {{eta*}}
and {{eps*}} using U-shaped connectors. In later diagrams, the wires
for {{1}} will be dropped as they can freely introduced or eliminated
using {{uniti}} and {{unite}}.

@@

%% {{(

@@

To build these isomorphisms, we first build an intermeidate
construction which we call {{eps_{fst} : (-b1)*b2 + b1*b2 <-> 0}}. 
\begin{center}
\scalebox{0.8}{
  \includegraphics{diagrams/eps_fst.pdf}
}
\end{center}

The isomorphism {{(-b1) * b2 <-> -(b1 * b2)}} can be constructed in
terms of {{eps_{fst} }} as shown below.

@@

The second isomorphism can be built in the same way by merely swapping
the arguments.

@@

This isomorphism can be built using \emph{involution} constructed in
Sec \ref{sec:neg} and the isomorphisms above, and correspond to the
algebraic manipulation:

%subcode{opsem} include main
%! columnStyle = rc
%   & b1 * b2
% = & -(-(b1*b2)) 
% = & -((-b1)*b2)
% = & (-b1)*(-b2)

@@

\paragraph*{QC} 
A \emph{time traveling} intuition is also applicable here.  In a
normal cicruit, as computational steps are taken values flow from left
to right, i.e. computational time progresses from left to right.  The
action of {{eps+}} causes values, aka information particles, to flow
from right to left i.e. backwards in computational time.  The
interesting thing about this interpretation is that particles that
travel backwards in time get to see and interact with the history of
particles that they coexist with. This gives us an intuitive
interpretation of the isomorphism {{(-b1)*b2<->-(b1*b2)}} (see
Sec. \ref{sec:specific-constructions}) where it would otherwise seem
that {{-b1}} and {{b2}} move in opposite directions. The backward flow
of the {{-b1}} allows it to `see the past' of {{b2}} and is thus
equivalent to both particles moving backward in time.

An analogy inspired by quantum mechanics is also applicable here.  The
operator {{eta*}} is a site of fission that creates \emph{an entangled
  particle and anti-particle}. These particles can be thought to be in
a \emph{superposition} of states determined by their type. Both flow
in the same direction in time.  In each possible world that the
particle exists, it takes on a specific value inhibiting its type and
correspondingly its entangled anti-particle takes the dual of the
specific value. Since the particles are entangled, actions on one
(such as transformation by application of an isomorphism) affects the
other in much the same sense as \emph{action at a distance}. The
`{{a-o*b}}' functions mentioned in the introduction, are first-class
values and they correspond to these entagled pairs. Finally {{eps*}}
operations are sites where corresponding particle and anti-particles
annihilate each other. While appealing as an analogy, this description
does not imply (or preclude) any formal connection with Physics.

Are constructions are very similar to those developed in the context
of QC. Is this how nature computes? Are the analogies to time
traveling particles and anti-particles more concrete in some way?
